---
title: "Week 16 - Model script"
author: "Amy Atkinson" 
date: "11/02/2022"
output: html_document
---

## The Wilcoxon rank-sum test

Background: You are a researcher interested in whether the old saying "an apple a day keeps the doctor away is true". You recruit 16 people and assign each participant to either a "0 apples" or "1 apple" condition. Participants in the "0 apple"" condition eat 0 apples every day for a year. Participants in the "1 apple" condition eat 1 apple a day for a year. You ask participants to report how many times they visited the GP in the year.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load in the libraries you need here
library(tidyverse) 
library(cowplot)
```

### Read in the data

```{r}
apple_data <- read.csv("apple_study_data.csv")
```

### Check whether the data meets the assumption of normality

#### 0 apples group:

##### Filter data:

This produces a new dataset containing data for the 0 apple a day group only:

```{r}
zero_apples <- apple_data %>% filter(Apples_per_day == "0 apples")  
```

##### Produce a Q-Q plot:

```{r}
qqnorm(zero_apples$Visits) # plots observed data
qqline(zero_apples$Visits, col = "steelblue") # plots expected data if data follows a normal distribution
```

Some evidence of deviation. 

##### Shapiro-Wilk test:


```{r}
shapiro.test(zero_apples$Visits)

```

This suggests that the data in the zero apples group violates the assumption of normality.

#### 1 apple group:

##### Filter data:

This produces a new dataset containing data for the 1 apple a day group only:

```{r}
one_apple <- apple_data %>% filter(Apples_per_day == "1 apple")

```

##### Produce a Q-Q plot:


```{r}
qqnorm(one_apple$Visits) # qqnorm plots our data

qqline(one_apple$Visits, col = "steelblue") # plots expected data if the data is from a normally distributed population.
```

The data in this group seems to follow the diagonal line quite well. 

##### Shapiro-Wilk test:

```{r}
shapiro.test(one_apple$Visits)
```

The Shapiro-Wilk test suggests the assumption of normality is not violated for this group

### Exploring our dataset:

Medians, range, and number of participants per group:

```{r}
apple_data %>% # 
  group_by(Apples_per_day) %>% # plots data per group
  summarise(med = median(Visits), # median Visits value 
            min = min(Visits), # lowest Visits value
            max = max(Visits), # highest Visits value
            n = n()) # number of participants per group
```

Plot:

```{r}
ggplot(apple_data, aes(x=Apples_per_day, y=Visits)) + 
  geom_boxplot() + 
  labs(y = "Number of visits to the GP") + 
  theme_cowplot() 
```

The median for the 1 apple group is lower than the median for the 0 apple group. There are 8 participants in each group.

### Performing the Wilcoxon rank-sum test:

```{r}
model1 <- wilcox.test(Visits ~ Apples_per_day, data = apple_data, paired = FALSE) 
model1
```

There is a significant difference between the groups (p = .018).

### Effect size: 

```{r}
z <- qnorm(model1$p.value/2)
r <- z/sqrt(16) # earlier, we saw we have 8 participants in each Apples_per_day, so 16
r

```

The effect size (r) is equal to -0.59. This is a large effect size.  


### Interpretation:

A Wilcoxon rank-sum test revealed that the participants who ate one apple a day (Median = 3.5; Range = 0-8) visited the GP less than participants who ate zero apples a day (Median = 9.5; Range = 2-41), W = 55, p = .018, r = -.59. 

This study provides evidence that an apple a day  keep the doctor away.

## The Wilcoxon signed-rank test

You are interested in whether eating bananas keeps the doctor away. This time you recruit only one group of participants. In the first year, you ask them to eat 0 banana every day. In the second year, you ask them to eat 1 bananas a day. You ask them to report how many times they visit the GP in Year 1 and Year 2. 

### Read in the data

This chunk reads in the data.

```{r}
banana_data <- read.csv("banana_study_data.csv") 
```

### Check whether the data meets the assumption of normality

```{r}
banana_data$difference <- banana_data$zero_banana - banana_data$one_banana 
```

#### Produce the Q-Q plot:

```{r}
qqnorm(banana_data$difference) # qqnorm plots our data

qqline(banana_data$difference, col = "steelblue") # plots expected data if the data is from a normally distributed population.
```

### Shapiro-Wilk test:

A few of the points deviate from the line. This suggests the assumption of normality may be violated.

```{r}
shapiro.test(banana_data$difference)

```

The p-value is significant, this suggests the assumption of normality has been violated. 

### Exploring the data:

#### Descriptive statistics:

```{r}
banana_data %>%
  summarise(med_0banana = median(zero_banana), min_0banana = min(zero_banana), max_0banana = max(zero_banana), # median, min and max for 0 banana timepoint,
            med_1banana = median(one_banana), min_1banana = min(one_banana), max_1banana = max(one_banana), # median, min and max for 1 banana timepoint
            n()) # number of participants
```
            
#### Plot:

```{r include =FALSE} 
# WE NEED THE DATA TO BE IN LONG FORMAT TO PLOT THE FIGURE
# set this to false if we don't want our html file to contain this chunk of code
banana_data_long <- gather(banana_data, Bananas, Visits, zero_banana:one_banana)
banana_data_long 

banana_data_long$Bananas <- factor(banana_data_long$Bananas, c("zero_banana", "one_banana")) # this will mean zero bananas is presented first on the olot
```

```{r}
ggplot(banana_data_long, aes(x=Bananas, y=Visits)) + 
  geom_boxplot() +
  labs(y = "Number of GP visits") +
  scale_x_discrete(labels = c("Zero", "One")) + # changes "zero_banana" to "Zero" and "one_banana" to "One" - you could also recode the variable to do the same thing
  theme_cowplot()
```

The median number of visits to the GP is lower in the zero banana condition than the one banana condition.

### Performing the Wilcoxon signed-rank test:

```{r}
model2 <- wilcox.test(banana_data$zero_banana, banana_data$one_banana, paired = TRUE)
model2
```

The p-value is equal to 0.03552 (or .036), indicating there is a significant difference between the conditions.

### Effect size: 

```{r} 
z_bananas <- qnorm(model2$p.value/2)
r_banana <- z_bananas/sqrt(12) # 12 because we have 12 observations overall (6 participants and 2 timepoints for each)
r_banana
```

The effect size (r) is equal to -0.61, indicating there is a large effect size.

### Interpretation:

A Wilcoxon signed-rank test revealed that the number of visits to the GP was significantly lower in the 0 banana condition (Median = 5; Range = 1-22) than the 1 apple a day condition (Median = 28; Range = 3-65), V = 0, p = .036, r = -.61. 

Given that participants who ate one banana a day visited the doctor more than participants who ate zero apples a day, this study provides evidence that eating a banana a day does not keep the doctor away 

(Disclaimer: data is 100% made up, please don't stop eating bananas based on this...!)
